---
title: "individual2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r}
HAPPY <- readRDS("C:/Users/Taylor Shaske/Downloads/HAPPY (1).rds")
data(happy, package="productplots")
head(happy)
str(happy)
HAPPY <- readRDS("C:/Users/Taylor Shaske/Downloads/HAPPY (1).rds")
knitr::opts_chunk$set(echo=TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(ggmosaic)
names(HAPPY)
```
'FINRELA':description of family income.
'WTSSALL':weight

#Cleaning
```{r}
HAPPY2 <- HAPPY
NAstrings <- c("IAP", "DK", "NA")
```
**HAPPY**
```{r}
is.na(HAPPY2$HAPPY) <- HAPPY$HAPPY %in% NAstrings
class(HAPPY$HAPPY)
HAPPY2$HAPPY <- factor(HAPPY2$HAPPY)
class(HAPPY2$HAPPY)
levels(HAPPY2$HAPPY)
ggplot(data = HAPPY2, aes(x=HAPPY)) + geom_bar()
```
For the factor `HAPPY`, "PRETTY HAPPY," had the highest level frequency, followed by "VERY HAPPY," "NOT TOO HAPPY," and then "NA".
**YEAR**
```{r}
is.na(HAPPY2$YEAR) <- HAPPY$YEAR %in% NAstrings
class(HAPPY$YEAR)
# don't need to change, it's already a factor variable
ggplot(data = HAPPY2, aes(x=YEAR)) + geom_histogram(bins=20)
```
There doesn't look like there is a trend for the number of observations in the data set per year. 
**AGE**
```{r}
is.na(HAPPY2$AGE) <- HAPPY$AGE %in% NAstrings
class(HAPPY$AGE)
HAPPY2$AGE <- factor(HAPPY2$AGE)
HAPPY2 <- HAPPY2 %>% mutate(AGE = recode(AGE, "89 OR OLDER" = "89"))
HAPPY2$AGE <- as.numeric(HAPPY2$AGE)
class(HAPPY2$AGE)
ggplot(HAPPY2, aes(x=AGE)) + geom_histogram(bins=15)
```
Age 16 has the highest number of observations. The histogram is skewed right so this means that the number of observations decreases as the `AGE` increases.
```{r}
HAPPY <- readRDS("C:/Users/Taylor Shaske/Downloads/HAPPY (1).rds")
names(HAPPY)
```
**SEX**
```{r}
is.na(HAPPY2$SEX) <- HAPPY$SEX %in% NAstrings
class(HAPPY$SEX)
HAPPY2$SEX <- factor(HAPPY$SEX)
class(HAPPY2$SEX)
levels(HAPPY2$SEX)
ggplot(data=HAPPY2, aes(x=SEX)) + geom_bar()
```
The data set shows that there were more "FEMALE" than "MALE" respondents. 
**MARITAL**
```{r}
is.na(HAPPY2$MARITAL) <- HAPPY$MARITAL %in% NAstrings
class(HAPPY$MARITAL)
HAPPY2$MARITAL <- factor(HAPPY$MARITAL)
class(HAPPY2$MARITAL)
levels(HAPPY2$MARITAL)
HAPPY2$MARITAL <- factor(HAPPY2$MARITAL, levels(HAPPY2$MARITAL)[c(3,2,4,1,5)])
levels(HAPPY2$MARITAL)
ggplot(data=HAPPY2, aes(x=MARITAL)) + geom_bar()
```
The `MARITAL` status with the most counts is "MARRIED," then "NEVER MARRIED," "DIVORCED," "WIDOWED," "SEPARATED," and "NA."
**DEGREE**
```{r}
is.na(HAPPY2$DEGREE) <- HAPPY$DEGREE %in% NAstrings
class(HAPPY$DEGREE)
HAPPY2$DEGREE <- factor(HAPPY2$DEGREE)
class(HAPPY2$DEGREE)
levels(HAPPY2$DEGREE)
HAPPY2$DEGREE <- factor(HAPPY2$DEGREE, levels(HAPPY2$DEGREE)[c(5,3,4,1,2)])
levels(HAPPY2$DEGREE)
ggplot(data=HAPPY2, aes(x=DEGREE)) + geom_bar()
```
Most people who responded have a "HIGH SCHOOL" degree, then the next highest is "LT HIGH SCHOOL," then "BACHELOR," then "GRADUATE," then "JUNIOR COLLEGE," and "NA".
**FINRELA**
```{r}
is.na(HAPPY2$FINRELA) <- HAPPY$FINRELA %in% NAstrings
class(HAPPY$FINRELA)
HAPPY2$FINRELA <- factor(HAPPY2$FINRELA)
class(HAPPY2$FINRELA)
levels(HAPPY2$FINRELA)
HAPPY2$FINRELA <- factor(HAPPY2$FINRELA, levels(HAPPY2$FINRELA)[c(5,3,2,1,4)])
levels(HAPPY2$FINRELA)
ggplot(data=HAPPY2, aes(x=FINRELA)) + geom_bar() + coord_flip()
```
We can tell that most people responded as "AVERAGE" for `FINRELA`. "BELOW AVERAGE" was more than "ABOVE AVERAGE" and more reported "FAR BELOW AVERAGE" than reported "FAR ABOVE AVERAGE." The amount of "NA" values is also higher than the number for "FAR BELOW AVERAGE" and "FAR ABOVE AVERAGE."
**HEALTH**
```{r}
is.na(HAPPY2$HEALTH) <- HAPPY$HEALTH %in% NAstrings
class(HAPPY$HEALTH)
HAPPY2$HEALTH <- factor(HAPPY2$HEALTH)
class(HAPPY2$HEALTH)
levels(HAPPY2$HEALTH)
HAPPY2$HEALTH <- factor(HAPPY2$HEALTH, levels(HAPPY2$HEALTH)[c(4,2,3,1)])
levels(HAPPY2$HEALTH)
ggplot(data=HAPPY2, aes(x=HEALTH)) + geom_bar()
```
The most commonly reported level for `HEALTH` was "GOOD," followed by "NA," "EXCELLENT," "FAIR," and "POOR".
**WTSSALL**
```{r}
is.na(HAPPY2$WTSSALL) <- HAPPY2$WTSSALL %in% NAstrings
class(HAPPY2$WTSSALL)
ggplot(data=HAPPY2, aes(x=WTSSALL)) + geom_histogram(bins=30)
```
The distribution of the `WTSSALL` count is skewed to the right..
**PARTYID**
```{r}
is.na(HAPPY2$PARTYID) <- HAPPY$PARTYID %in% NAstrings
class(HAPPY$PARTYID)
HAPPY2$PARTYID <- factor(HAPPY2$PARTYID)
class(HAPPY2$PARTYID)
levels(HAPPY2$PARTYID)
HAPPY2$PARTYID <- factor(HAPPY2$PARTYID, levels(HAPPY2$PARTYID)[c(7,4,1,3,2,5,8,6)])
levels(HAPPY2$PARTYID)
ggplot(data=HAPPY2, aes(x=PARTYID)) + geom_bar() + coord_flip()
```
For the `PARTYID` data set, "NOT STR DEMOCRAT" had the highest count, then it was "STRONG DEMOCRAT," "NOT STR REPUBLICAN," "INDEPENDENT," "IND,NEAR DEM," "STRONG REPUBLICAN," "IND,NEAR REP," "OTHER PARTY," and "NA".
**POLVIEWS**
```{r}
is.na(HAPPY2$POLVIEWS) <- HAPPY$POLVIEWS %in% NAstrings
class(HAPPY$POLVIEWS)
HAPPY2$POLVIEWS <- factor(HAPPY2$POLVIEWS)
class(HAPPY2$POLVIEWS)
levels(HAPPY2$POLVIEWS)
HAPPY2$POLVIEWS <- factor(HAPPY2$POLVIEWS, levels(HAPPY2$POLVIEWS)[c(2,4,7,5,6,1,3)])
levels(HAPPY2$POLVIEWS)
ggplot(data=HAPPY2, aes(x=POLVIEWS)) + geom_bar() + coord_flip()
```
"MODERATE" was most common by far, then it was "NA," "SLGHTLY CONSERVATIVE," "CONSERVATIVE," "SLIGHTLY LIBERAL," "LIBERAL," "EXTRMLY CONSERVATIVE," and "EXTREMELY LIBERAL."
#Data Exploration
**Distribution of `MARITAL`**
```{r}
data <- HAPPY[!is.na(HAPPY$MARITAL),]
ggplot(data, aes(x=MARITAL)) + geom_bar(width=1)
#Rearrange bars in order of frequency and remove NA values. 
noNAMAR <- HAPPY2 %>% filter(!is.na(MARITAL))
marplot1 <- ggplot(mutate(noNAMAR, MARITAL =(MARITAL))) + geom_bar(aes(x = MARITAL)) + xlab("MARITAL") + ggtitle("MARITAL Distribution")
marplot1
#Look at distribution of `MARITAL` based on age.
marplot2 <- ggplot(noNAMAR, aes(x=MARITAL, y=AGE)) + geom_boxplot() + ggtitle("MARITAL by AGE")
marplot2
```
The distribution of `MARITAL` appears to relate to `AGE`. The boxplot shows that "widowed" has the highest interquartile numbers and median, while never married has the lowest for both.
**Distribution of `HEALTH`**
```{r}
ggplot(data=HAPPY2, aes(x=HEALTH)) + geom_bar()
noNAHEALTH <- HAPPY2 %>% filter(!is.na(HEALTH))
plot1 <- ggplot(mutate(noNAHEALTH, HEALTH =(HEALTH))) + geom_bar(aes(x = HEALTH)) + xlab("HEALTH") + ggtitle("HEALTH Distribution")
plot1
plot2 <- ggplot(noNAHEALTH, aes(x=HEALTH, y=AGE)) + geom_boxplot() + ggtitle("HEALTH by AGE")
plot2
```

This boxplot shows us that HEALTH is connected to your age. As expected the median for poor health has a higher age than the rest of them, while the median for excellent health has the lowest age.
**`HAPPY` and ` AGE`**
```{r}
ggplot(HAPPY2, aes(x=HAPPY, y=AGE)) + geom_boxplot() + ggtitle("HAPPY by AGE")
noNAHAPPY <- HAPPY2 %>% filter(!is.na(HAPPY))
ggplot(noNAHAPPY, aes(x=HAPPY, y=AGE)) + geom_boxplot() + ggtitle("HAPPY by AGE")
```
The boxplots for HAPPY by AGE are too close for one to affect the other, so we can conclude that happiness age doesn't affect your happiness.
**`HAPPY` and `POLVIEWS`**
```{r}
noNAHAPPY <- noNAHAPPY %>% filter(!is.na(POLVIEWS))
ggplot(noNAHAPPY) + geom_mosaic(aes(x=product(HAPPY), fill=POLVIEWS)) + ggtitle("HAPPY by POLVIEWS") + ylab("POLVIEWS") + xlab("HAPPY")
```

**`HAPPY` and `SEX`**
```{r}
ggplot(noNAHAPPY) + geom_mosaic(aes(x=product(SEX), fill=HAPPY)) + ggtitle("HAPPY by SEX") + ylab("HAPPY") + xlab("SEX")
#ggplot(HAPPY2) + geom_mosaic(aes(x=product(AGE), fill=HAPPY))
#ggplot(data=HAPPY2, aes(x=HAPPY, y=AGE)) + geom_point()
#ggplot(HAPPY2) + geom_mosaic(aes(x=product(HAPPY), fill=DEGREE)) + facet_grid(SEX ~ MARITAL)
```
It doesn't look like one `SEX` is much different in happiness than the other. `SEX` does not affect `HAPPY` according to the plot.